---
title: "USED CAR PRICES CASE STUDY"
subtitle: 'Deliverable I: Data Processing, Description, Validation and Profiling'
author: "Miquel Parra i Xavier Alaman"
date: \today
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

# R libraries imports, useful functions and data loading

In this first section we will load all required packages and libraries, declare additional functions, and load our data.

## Load Required Packages

```{r}
# Load Required Packages: to be increased over the course
options(contrasts=c("contr.treatment","contr.treatment"))

requiredPackages <- c("effects","FactoMineR","car","missMDA","mvoutlier","chemometrics", "factoextra","RColorBrewer","ggplot2","dplyr","ggmap","ggthemes","knitr")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(requiredPackages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
#verify they are loaded
search()

```

## Sample load

```{r, results='hide'}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())

# Users file path
miquel_fp <- "C:/Users/Miquel/Documents/GitHub/ADEI/"
xavi_fp <- "~/Documents/FIB/ADEI/ADEI/"
filepath <- xavi_fp
filepath <-miquel_fp

# Set working directory
setwd(filepath)

# Load data from file
load(paste0(filepath,"MyOldCars-5000Clean.RData"))

# Index reset
row.names(df) <- NULL
```

## PCA

We have done PCA in order to combine related variables, and focus on uncorrelated or independent ones, especially those along which the observations have high variance. Also, to obtain a smaller set of variables that explain most of the variance in the original data, in more compact and insightful form.
```{r}
names(df)
vars_con<-names(df)[c(5,7,8,11)]
vars_res<-names(df)[c(3,18)]
vars_dis<-names(df)[c(1,2,4,6,9,10,13,14,15,16,17)]

vars_con
vars_res
vars_dis

# Multivariant outliers should be included as supplementary observations
ll <- which( df$mout == "YesMOut")
res.pca<-PCA(df[,c(vars_res, vars_dis, vars_con)],quali.sup=c(2:13),quanti.sup= c(1), ind.sup = ll ) 

summary(res.pca)
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 60))

sort(res.pca$ind$contrib[,1],decreasing=T)[1:6]
sort(res.pca$ind$contrib[,2],decreasing=T)[1:6]
sort(res.pca$ind$contrib[,1:2],decreasing=T)[1:6]

sort(res.pca$var$contrib[,1],decreasing=T)[1:3]
sort(res.pca$var$cos2[,1],decreasing=T)[1:3]
sort(res.pca$var$coord[,1],decreasing=T)[1:4]
sort(res.pca$var$contrib[,2],decreasing=T)[1:3]
sort(res.pca$var$cos2[,2],decreasing=T)[1:3]
sort(res.pca$var$coord[,2],decreasing=T)[1:4]

fviz_contrib(res.pca, choice = "ind", axes = 1:2, top = 20)
fviz_contrib(res.pca, choice = "var", axes = 1:2)
fviz_cos2(res.pca, choice = "var", axes = 1:2)
```
According to Kaiser criteria (principal components with an eigenvalue > 1) we have to interpret two axes (Dim.1 and Dim.2).
According to Elbow's rule (based on selecting dimensions until the difference in variance of that of the next factorial plane is almost the same as that of the current plane) we have to interpret two axes (Dim.1 and Dim.2).
The 6 most contributive individuals for 1 principal component axis in descending order are 2323, 2074, 1027, 488, 1138 and 3933.
The 6 most contributive individuals for 2 principal component axis in descending order are 2703, 4570, 2429, 1839, 239 and 324. 

The 3 most contributive variables for 1 principal component axis in descending order are years_sell, mileage and mpg. 
The 3 most contributive variables for 2 principal component axis in descending order are tax, mpg and years_sell.

The 3 best represented variables for 1 principal component axis in descending order are years_sell, mileage and mpg. 
The 3 best represented variables for 2 principal component axis in descending order are tax, mpg and years_sell.

Better explanation I think:
As we can see from the Contribution of individuals to Dim-1-2 plot, the 20 most contributive individuals in descending order are 4570, 2703, 3254, 2199, 4662, 735, 4566, 1944, 99, 913, 4997, 80, 224, 2429, 2353, 1839, 239, 4596, 531 and 324. 
As we can see from the Contribution of variables to Dim-1-2 plot, the 4 most contributive variables in descending order are years_sell, mileage, tax and mpg. 
As we can see from the Cos2 of variables to Dim-1-2, the 4 best represented variables in descending order are years_sell, mileage, tax and mpg. 

The years_sell, mileage, mpg and tax variables have a positive first dimension.
The tax, years_sell and mileage variables have a positive second dimension and the mpg variable has a negative second dimension. 

Dimension 1 shows a positive correlation to years_sell and mileage. Price target variable (supplementary) is negatively associated to the first factorial axis. 
Dimension 2 shows the greatest positive correlation to tax and the lowest negative correlation to mpg. 

##Hierarchical Clustering
```{r}
# 2 dimensions have to be selected according to Kaiser's criteria
res.pca<-PCA(df[,c(vars_res, vars_dis, vars_con)],quali.sup=c(2:13),quanti.sup= c(1), ind.sup = ll,ncp =2 )
res.hcpc<-HCPC(res.pca,nb.clust=-1,graph=T)
res.hcpc$desc.var

df$claHC<-4
df[ row.names(res.hcpc$data.clust), "claHC" ]<-res.hcpc$data.clust$clust
df$claHC<-factor(df$claHC)
levels( df$claHC ) <- paste0("claHCPC-",levels( df$claHC ) )
```
Cluster 1: As we can see price mean is greater in cluster 1 than in the overall mean. Also, we can see that tax, mpg, mileage and years_sell means are lower in cluster 1 than in the overall means. 
So we can say that cluster 1 is the cluster of expensive cars with few/less tax, mpg, mileage and years_sell.

Cluster 2: As we can see tax, years_sell and mileage means are greater in cluster 2 than in the overall means. Also, we can see that mpg mean in cluster 2 is lower than in the overall mean. 
So we can say that cluster 2 is the cluster of cars with high/great tax, years_sell and mileage and with few/less mpg. 

Cluster 3: As we can see years_sell, mileage and mpg means are greater in cluster 3 than in the overall means. Also, we can see that tax and price means are lower in cluster 3 than in the overall means. 
So we can say that cluster 3 is the cluster of cheap cars with high/great years_sell, mileage and mpg and with less/few tax. 

Cluster 4: Cluster of multivariant outliers. 

##K-Means Classification
```{r}
res.pca<-PCA(df[,c(vars_res, vars_dis, vars_con)],quali.sup=c(2:13),quanti.sup= c(1), ind.sup = ll,ncp =2 )
ppcc<-res.pca$ind$coord[,1:2]
dim(ppcc)

dist<-dist(ppcc)
kc<-kmeans(dist,3, iter.max = 30, trace=T)

df$claKM<-4
df[ -ll, "claKM" ]<-kc$cluster
df$claKM<-factor(df$claKM)
# levels( df$claKM ) <- paste0("claKMPC-",levels( df$claKM ) )

kc$betweenss/kc$totss

names(df)
catdes(df,21)

```
Cluster 1: As we can see price mean is greater in cluster 1 than in the overall mean. Also, we can see that tax, mpg, mileage and years_sell means are lower in cluster 1 than in the overall means. So we can say that cluster 1 is the cluster of expensive cars with few/less tax, mpg, mileage and years_sell.(totalMOE negative)

Cluster 2: As we can see tax, years_sell and mileage means are greater in cluster 2 than in the overall means. Also, we can see that price and mpg means are lower in cluster 2 than in the overall means. So we can say that cluster 2 is the cluster of cars with high/great tax, years_sell and mileage and with few/less price and mpg. (totalMOE negative)

Cluster 3: As we can see mpg, years_sell and mileage means are greater in cluster 3 than in the overall means. Also, we can see that tax and price means are lower in cluster 3 than in the overall means. So we can say that cluster 3 is the cluster of cars with high/great mpg, years_sell and mileage and with few/less tax and price.(totalMOE positive)

Cluster 4: Cluster of multivariant outliers
As we can see mileage, years_sell and tax means are greater in cluster 4 than in the overall means. Also, we can see that mpg mean is lower in cluster 4 than in the overall mean. So we can say that cluster 4 is the cluster of cars with high/great mileage, years_sell and tax and with few/less mpg. (totalMOE positive)

##CA analysis for your data should contain your factor version of the numeric target (previous) in K= 7 (maximum 10) levels and 2 factors:
```{r}
par(mfrow=c(1,1))

df[,c(13,15)]

tt<-table(df[,c("aux_price","aux_tax")])
tt
res.ca<-CA(tt)
# lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
chisq.test(tt)

names(res.ca)
fviz_eig(res.ca) #Same outputs as PCA
summary(res.ca,dig=2)

plot( res.ca, cex=0.8, graph.type = "classic" )
lines( res.ca$row$coord[,1], res.ca$row$coord[,2], col="blue", lwd = 2 )
lines( res.ca$col$coord[,1], res.ca$col$coord[,2], col="red", lwd = 2 )
fviz_ca_biplot(res.ca,repel=TRUE)+theme_bw() 
```
As we can see aux_price and aux_tax variables are not independent because the p-value is less than 0.05.
Also, we can see that the cheapest cars have the highest taxes and the most expensive cars have average/medium taxes.


We have considered to impute hybrid cars because they cause noise.//Revisar
```{r}
par(mfrow=c(1,1))
#lh<-which(df$fuelTyp=="f.Fuel-Hybrid")
#df<-df[-lh,]
tt<-table(df[,c("aux_price","fuelType")])
tt
res.ca<-CA(tt)
plot( res.ca, cex=0.8, graph.type = "classic" )
lines( res.ca$row$coord[,1], res.ca$row$coord[,2], col="blue", lwd = 2 )
lines( res.ca$col$coord[,1], res.ca$col$coord[,2], col="red", lwd = 2 )
summary(res.ca)
chisq.test(tt)
```
As we can see aux_price and aux_tax variables are not independent because the p-value is less than 0.05.
Also, we can see that hybrid cars cause noise. We can see that diesel cars have the highest prices and petrol and other cars have the lowest prices. 

The objectives of MCA are close to PCA objectives: similarity among individuals/ similarity among variables. Also, generalization of CA objectives: relationships among the categories.
##MCA analysis for your data should contain:
```{r}
par(mfrow=c(1,1))
llvout<-which(df$mout=="YesMOut");length(llvout) #Multivariate outliers
res.mca<-MCA(df[,c(vars_dis[c(3:4,6:11)],"price", "Audi") ],quali.sup=c(4,10),quanti.sup=9 , ind.sup=llvout)
mean(res.mca$eig[,1])
summary(res.mca)
vars_dis
fviz_eig(res.mca)

sort(res.mca$ind$contrib[,1],decreasing=T)[1:6]
sort(res.mca$ind$contrib[,2],decreasing=T)[1:6]
sort(res.mca$ind$contrib[,3],decreasing=T)[1:6]
sort(res.mca$ind$contrib[,4],decreasing=T)[1:6]
sort(res.mca$ind$contrib[,5],decreasing=T)[1:6]
#sort(res.mca$ind$contrib[,6],decreasing=T)[1:6]
#sort(res.mca$ind$contrib[,7],decreasing=T)[1:6]
#sort(res.mca$ind$contrib[,8],decreasing=T)[1:6]
#sort(res.mca$ind$contrib[,9],decreasing=T)[1:6]

fviz_contrib(res.mca, choice = "ind", axes = 1:5, top = 20)#NOOOOOO EM DEIXA POSAR 1:9
vars_dis
```
According to Kaiser criteria (principal components with an eigenvalue > mean(eigenvalues)) we have to interpret 9 axes (Dim. 1 to Dim. 9).

The 6 most contributive individuals for 1 principal component axis in descending order are 1282, 1646, 1900, 1980, 2047 and 2048. 
The 6 most contributive individuals for 2 principal component axis in descending order are 3537, 3650, 3719, 3761, 3993 and 4543.
The 6 most contributive individuals for 3 principal component axis in descending order are 493, 851, 913, 1863, 425 and 492.
The 6 most contributive individuals for 4 principal component axis in descending order are 1529, 1544, 1866, 1998, 2013 and 1108.
The 6 most contributive individuals for 5 principal component axis in descending order are 1102, 1210, 1626, 1663, 1207 and 1401. 
The 6 most contributive individuals for 6 principal component axis in descending order are 
The 6 most contributive individuals for 7 principal component axis in descending order are 
The 6 most contributive individuals for 8 principal component axis in descending order are 
The 6 most contributive individuals for 9 principal component axis in descending order are 

Better explanation I think:
As we can see from the Contribution of individuals to Dim-1-5 plot, the 20 most contributive individuals in descending order are 1102, 1210, 1626, 1663, 1207, 1401, 1639, 1696, 1295, 1684, 1798, 1763, 1891, 1280, 1487, 1636, 1993, 1999, 1912 and 1530. 

We can see/distinguish four clouds of variables (variables in the same cloud means they are related):
First: manufacturer, transmission and fuelType variables.
Second: Audi variable. 
Third: aux_price, price, aux_mpg and aux_tax variables. 
Forth: aux_mileage and aux_years_sell variables. 

As we can see the cheapest cars are VW, have manual transmission, high taxes, medium years_sell, medium-high mileage and medium-high mpg.
As we can see the most expensive cars are Mercedes, hybrid, have automatic or semi-automatic transmission, medium taxes, few years_sell, low mileage and low mpg. 

As we can see Audi cars are petrol or other and have medium-low mpg. 

As we can see supplementary variable price is in the second quadrant which corresponds to the most expensive cars. 


##Hierarchical Clustering (from MCA)
```{r}
#5 dimensions 
res.mca<-MCA(df[,c(vars_dis[c(3:4,6:11)],"price") ],quali.sup=c(4,6),quanti.sup=9 , ind.sup=llvout, ncp = 5)
res.hcmc<-HCPC(res.mca,nb.clust=-1,order=TRUE)
names(res.hcmc)

res.hcmc$desc.var

df$clamHC<-4
df[ row.names(res.hcmc$data.clust), "clamHC" ]<-res.hcmc$data.clust$clust
df$clamHC<-factor(df$clamHC)
levels( df$clamHC ) <- paste0("claHCMC-",levels( df$clamHC ) )

res.hcmc$desc.ind$para
res.hcmc$desc.ind$dist

table(df$clamHC,df$claHC)
table(df$clamHC,df$claKM)

```
Cluster 1: We can say that cluster 1 is the cluster of expensive cars (overrepresented) with low mileage (overrepresented), low mpg (overrepresented), few years_sell (overrepresented) and medium taxes (overrepresented).

Cluster 2: We can say that cluster 2 is the cluster of cheap cars (overrepresented) with low taxes (overrepresented), medium-high mileage (overrepresented), medium years_sell (overrepresented) and high mpg (overrepresented).

Cluster 3: We can say that cluster 3 is the cluster of cheap cars (overrepresented) with high years_sell (overrepresented), high mileage (overrepresented), high mpg (overrepresented) and high taxes (overrepresented).  

Cluster 4: Cluster of multivariant outliers.

The parangons of cluster 1 are 4080, 4572, 4668, 4811 and 4837.
The parangons of cluster 2 are 4562, 4991, 4912, 140 and 610.
The parangons of cluster 3 are 3558, 4083, 4988, 3935 and 3962.

The dist/distants of cluster 1 are 1102, 1210, 1626, 1663 and 1207.
The dist/distants of cluster 2 are 1530, 1989, 2023, 1280 and 1487.
The dist/distants of cluster 3 are 2547, 3295, 469, 1807 and 1829.

