---
title: "Load Data and Set Sample"
author: "Josep Franquet Fabregas"
date: \today
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    toc: no
    toc_depth: '4'
  word_document:
    toc: no
    toc_depth: '4'
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
subtitle: 'Laboratori 1 - Data Preparation'
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

# USED CAR PRICES CASE STUDY -- DELIVERABLE 1

## Data Description: 100,000 UK Used Car Data set

This data dictionary describes data (<https://www.kaggle.com/adityadesai13/used-car-dataset-ford-and-mercedes>) - A sample of 5000 trips has been randomly selected from Mercedes, BMW, Volkwagen and Audi manufacturers. So, firstly you have to combine used car from the 4 manufacturers into 1 dataframe.

The cars with engine size 0 are in fact electric cars, nevertheless Mercedes C class, and other given cars are not electric cars,so data imputation is requered.

-   manufacturer Factor: Audi, BMW, Mercedes or Volkswagen
-   model Car model
-   year registration year
-   price price in £
-   transmission type of gearbox
-   mileage distance used
-   fuelType engine fuel
-   tax road tax
-   mpg Consumption in miles per gallon\
-   engineSize size in litres

## Load Required Packages: to be increased over the course

```{r}
# Load Required Packages: to be increased over the course
options(contrasts=c("contr.treatment","contr.treatment"))

requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","dplyr","ggmap","ggthemes","knitr")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(requiredPackages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
#verify they are loaded
search()

```

## Load our sample

```{r}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())

setwd("C:/Users/Miquel/Documents/GitHub/ADEI")
filepath<-"C:/Users/Miquel/Documents/GitHub/ADEI/"

load(paste0(filepath,"MyOldCars-Raw.RData"))
```

## Some useful functions

```{r, echo=FALSE}

# Mout <- which((df$tax < var_out$mouti)|(df$tax > var_out$mouts))

# Some useful functions
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

```

## Validation of the Data Set: description of the process

### Univariate Descriptive Analysis 

#### Model variable
```{r}

df$model<-factor(paste0(df$manufacturer,"-",df$model))
df$model<-factor(df$model,labels=paste0("f.Model",levels(df$model)))

summary(df$model)
barplot(summary(df$model), main = "Model Barplot", col = "blue")

```

#### Year variable
```{r}
df$years_sell <-  2022 - df$year ####AQUI????????????????????????????????????????????????????????????????????????????????????
df$year <- factor(df$year)
df$year <- factor(df$year, labels = paste0("f.Year-",levels(df$year)))
summary(df$year)
barplot(summary(df$year), main = "Year Barplot", col = "blue")
```

#### Transmission variable
```{r}
df$transmission<-factor(df$transmission)
df$transmission <- factor(df$transmission, levels = c("Manual","Semi-Auto","Automatic"),labels = paste0("f.Trans-",c("Manual","SemiAuto","Automatic")))
summary(df$transmission)
barplot(summary(df$transmission), main = "Transmission Barplot", col = "blue")
```

#### fuelType variable
```{r}
df$fuelType<-factor(df$fuelType)
df$fuelType <- factor(df$fuelType, labels = paste0("f.Fuel-",levels(df$fuelType)))
summary(df$fuelType)
barplot(summary(df$fuelType), main = "FuelType Barplot", col = "blue")
```

#### engineSize variable
```{r}
df$engineSize<-factor(df$engineSize)
df$engineSize <- factor(df$engineSize, labels = paste0("f.EngSize-",levels(df$engineSize))) 
summary(df$engineSize)
barplot(summary(df$engineSize), main = "EngineSize Barplot", col = "blue")
```

#### manufacturer variable
```{r}
df$manufacturer<-factor(df$manufacturer)
df$manufacturer <- factor(df$manufacturer, labels = paste0("f.Mfr-",levels(df$manufacturer)))
summary(df$manufacturer)
barplot(summary(df$manufacturer), main = "Manufacturer Barplot", col = "blue")
```

#### price variable
```{r}
summary(df$price)
Boxplot(df$price)
df$f.price<-factor(cut(df$price/1000,breaks=c(quantile(df$price,seq(0,1,0.25),na.rm=TRUE))/1000,include.lowest = T ))
summary(df$f.price)####))/1000
levels(df$f.price)<-paste("f.price-",levels(df$f.price))
table(df$f.price,useNA="always")


```

#### mileage variable
```{r}
summary(df$mileage)
Boxplot(df$mileage)
df$f.mileage<-factor(cut(df$mileage/1000,breaks=c(quantile(df$mileage,seq(0,1,0.25),na.rm=TRUE))/1000,include.lowest = T ))
summary(df$f.mileage)
levels(df$f.mileage)<-paste("f.mileage-",levels(df$f.mileage))
table(df$f.mileage,useNA="always")
```

#### tax variable
```{r}
summary(df$tax)
Boxplot(df$tax)
df$f.tax<-factor(cut(df$tax,breaks=c(0,125,145,155,580),include.lowest = T ))
summary(df$f.tax)
levels(df$f.tax)<-paste("f.tax-",levels(df$f.tax))
table(df$f.tax,useNA="always")
```

#### mpg variable
```{r}
summary(df$mpg)
Boxplot(df$mpg)
df$f.mpg<-factor(cut(df$mpg,breaks=c(quantile(df$mpg,seq(0,1,0.25),na.rm=TRUE)),include.lowest = T ))
summary(df$f.mpg)
levels(df$f.mpg)<-paste("f.mpg-",levels(df$f.mpg))
table(df$f.mpg,useNA="always")
```

#### years_sell variable
```{r}
summary(df$years_sell)
Boxplot(df$years_sell)
df$f.years_sell<-factor(cut(df$years_sell,breaks=c(quantile(df$years_sell,seq(0,1,0.25),na.rm=TRUE)),include.lowest = T ))
summary(df$f.years_sell)
levels(df$f.years_sell)<-paste("f.years_sell-",levels(df$f.years_sell))
table(df$f.years_sell,useNA="always")
```

### Initialization of counts for missings, outliers and errors. All numerical variables have to be checked before

```{r}

#######################################################
imis<-rep(0,nrow(df))  # rows - cars
jmis<-rep(0,ncol(df))  # columns - variables
######################################################
mis1<-countNA(df) #There are no missings at the beginning 
mis1$mis_ind # Number of missings for the current set of cars
mis1$mis_col # Number of missings for the current set of variables

sum(mis1$mis_ind)
sum(mis1$mis_col)

#######################################################
iouts<-rep(0,nrow(df))  # rows - cars
jouts<-rep(0,ncol(df))  # columns - variables
######################################################

#######################################################
ierrs<-rep(0,nrow(df))  # rows - cars
jerrs<-rep(0,ncol(df))  # columns - variables
######################################################
#sobra 2 diu profe
```

### Data Quality Report
#### Missings
```{r}
mis1<-countNA(df)
mis1$mis_ind
mis1$mis_col
summary(mis1$mis_ind)
summary(mis1$mis_col)
```


#### engineSize variable
```{r}
sel<-which(df$engineSize==0 & df$fuelType!="f.Fuel-Electric")
ierrs[sel]<-ierrs[sel]+1
jerrs[9] <- length(sel)
```

#### price variable
```{r}
Boxplot(df$price)
var_out<-calcQ(df$price)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

# Outliers:
llout_price<-which((df$price<var_out$souti)|(df$price>var_out$souts))#souts abline
iouts[llout_price]<-iouts[llout_price]+1
jouts[3]<-length(llout_price)
```

#### mileage variable
```{r}
Boxplot(df$mileage)
var_out<-calcQ(df$mileage)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

# Outliers:
llout_mileage<-which((df$mileage<var_out$souti)|(df$mileage>var_out$souts))#souts abline
iouts[llout_mileage]<-iouts[llout_mileage]+1
jouts[5]<-length(llout_mileage)
# Assignation of an NA to my response variable:
df[llout_mileage,"mileage"]<-NA #llout
```

#### tax variable
```{r}
Boxplot(df$tax)
var_out<-calcQ(df$tax)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

# Outliers:
llout_tax<-which((df$tax<var_out$souti)|(df$tax>var_out$souts))#souts abline
iouts[llout_tax]<-iouts[llout_tax]+1
jouts[7]<-length(llout_tax)
# Assignation of an NA to my response variable:
df[llout_tax,"tax"]<-NA #llout
```

#### mpg variable
```{r}
Boxplot(df$mpg)
var_out<-calcQ(df$mpg)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

# Outliers:
llout_mpg<-which((df$mpg<var_out$souti)|(df$mpg>var_out$souts))#souts abline
iouts[llout_mpg]<-iouts[llout_mpg]+1
jouts[8]<-length(llout_mpg)
# Assignation of an NA to my response variable:
df[llout_mpg,"mpg"]<-NA #llout
```

#### years_sell variable
```{r}
Boxplot(df$years_sell)
var_out<-calcQ(df$years_sell)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

# Outliers:
llout_years_sell<-which((df$years_sell<var_out$souti)|(df$years_sell>var_out$souts))#souts abline
iouts[llout_years_sell]<-iouts[llout_years_sell]+1
jouts[11]<-length(llout_years_sell)
# Assignation of an NA to my response variable:
df[llout_years_sell,"years_sell"]<-NA #llout
```

#### Number of missing values of each variable
```{r}
missings_ranking_sortlist <- sort.list(mis1$mis_col, decreasing = TRUE)
for(j in missings_ranking_sortlist) {
  print(paste(names(df)[j]," : ", mis1$mis_col$mis_x[j]))
}
```

#### Number of errors per each variable
```{r}
errors_ranking_sortlist <- sort.list(jerrs, decreasing = TRUE)
for(j in errors_ranking_sortlist) {
  if(!is.na(names(df)[j])) { print(paste(names(df)[j], " : ", jerrs[j])) }
}
```

#### Number of outilers per each variable
```{r}
outliers_ranking_sortlist <- sort.list(jouts, decreasing = TRUE)
for(j in outliers_ranking_sortlist) {
  if(!is.na(names(df)[j])) print(paste(names(df)[j], " : ", jouts[j]))
}
```

#### Number of missing values individual
```{r}
barplot(table(imis),main = "Missings per individual Barplot", col = "blue")
```

#### Number of errors individual
```{r}
barplot(table(ierrs), main = "Errors per individual Barplot", col = "blue")
```

#### Number of outliers individual
```{r}
barplot(table(iouts), main = "Outliers per individual Barplot", col = "blue")
```

#### Multivariant Outlier detection 

```{r}
library(mvoutlier)
vars_con<-names(df)[c(3,5,7,8, 11)]
ll<-which(is.na(df$price))
if (length( ll ) > 0 ) df<-df[-ll,]
summary(df[,vars_con])
# aq.plot(df[,c(vars_con)]) # Might not work
names(df)
vars_con
names(df)
mout<-aq.plot(df[,vars_con],delta=qchisq(0.995,5),quan=0.995)

library(chemometrics)
summary(df[,vars_con])
mout<-Moutlier(df[,c(3,5,8)],quantile = 0.995, plot = TRUE)#FALTA LA 11 years_sell

ll<-which(mout$rd>mout$cutoff)
Boxplot(mout$rd)
df[ll,c(vars_res,vars_con)]
df$mout <- 0
df$mout[ ll ]<-1
df$mout <- factor( df$mout, labels=c( "NoMOut","YesMOut"))
table(df$mout)
```

#### Create variable adding the total number missing values, outliers and errors
```{r}
total_missings <- 0
total_errors <- 0
total_outliers <- 0

for(m in imis) { total_missings <- total_missings + m}
for(e in ierrs) {total_errors <- total_errors + e  }
for(o in iouts) {total_outliers <- total_outliers + o }
```

#### FALTAAAAAAAAAAAAAAAAAAAAAAAAAAAAA CORRELATION I AIXOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
```{r}
#CONSIDERAR AQUESTES AGRUPACIONS TMBBBBBBBBBBBBBBBBBBBB "VALORS RESIDUALS"
table(df$year)
df[which(df$year<=2009),"year"] <- "2009 or before"
df$year<-factor(df$year)
levels(df$year)
barplot(table(df$year))

table(df$engineSize)
df[which(df$engineSize>=4.0),"engineSize"] <- "4.0 or more" #or with 3.2 or 4.2

table(df$years_sell)
df[which(df$years_sell>=13),"years_sell"] <- "13 or more"
```



### Imputation numeric variables
```{r}
library(missMDA)
names(df)
vars_con<-names(df)[c(5,7,8, 11)] #Variable target NOOOOOOOOOOOOOOOOOOOOO
vars_res<-names(df)[c(3)]

summary(df[,vars_con])
res.impca<-imputePCA(df[,vars_con],ncp=4)

summary(res.impca$completeObs)
```


#### mileage variable
```{r}
par(nfrow = c(1,2))
hist(df$mileage)
hist(res.impca$completeObs[,"mileage"])
quantile(df$mileage,seq(0,1,0.1),na.rm=T)
round(quantile(res.input$completeObs[,"mileage"],seq(0,1,0.1),na.rm=T),dig=1)
```


#### tax variable
```{r}
par(nfrow=c(1,2))
hist(df$tax)
hist(res.impca$completeObs[,"tax"])
quantile(df$tax,seq(0,1,0.1),na.rm=T)
round(quantile(res.input$completeObs[,"tax"],seq(0,1,0.1),na.rm=T),dig=1)
```


#### mpg variable
```{r}
par(nfrow = c(1,2))
hist(df$mpg)
hist(res.impca$completeObs[,"mpg"])
quantile(df$mpg,seq(0,1,0.1),na.rm=T)
round(quantile(res.input$completeObs[,"mpg"],seq(0,1,0.1),na.rm=T),dig=1)
```


#### years_sell variable
```{r}
par(nfrow = c(1,2))
hist(df$years_sell)
hist(res.impca$completeObs[,"years_sell"])
quantile(df$years_sell,seq(0,1,0.1),na.rm=T)
round(quantile(res.input$completeObs[,"years_sell"],seq(0,1,0.1),na.rm=T),dig=1)
```

```{r}
df[ , vars_con ]<-res.impca$completeObs
```


### Imputation factors
```{r}
vars_dis<-names(df)[c(1:2, 4, 6, 9, 10)]
vars_res<-names(df)[c(3)]
summary(df[,vars_dis])
res.immca<-imputeMCA(df[,vars_dis],ncp=10)
summary(res.immca$completeObs)
df [c(1:2, 4, 6, 9, 10)] <- res.immca$completeObs
```


#### model variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$model),col="red")
barplot(table(res.input$completeObs[,"model"]),col="blue")
```


#### year variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$year),col="red")
barplot(table(res.input$completeObs[,"year"]),col="blue")
```


#### transmission variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$transmission),col="red")
barplot(table(res.input$completeObs[,"transmission"]),col="blue")
```


#### fuelType variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$fuelType),col="red")
barplot(table(res.input$completeObs[,"fuelType"]),col="blue")
```


#### engineSize variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$engineSize),col="red")
barplot(table(res.input$completeObs[,"engineSize"]),col="blue")
```


#### manufacturer variable
```{r}
par(mfrow=c(1,2))
barplot(table(df$manufacturer),col="red")
barplot(table(res.input$completeObs[,"manufacturer"]),col="blue")
```

```{r}
df[ , vars_dis ]<-res.immca$completeObs 
sum(countNA(df)$mis_ind)==0
```


### Profiling
```{r}
library(FactoMineR)
summary(df$price)
# The "variable to describe cannot have NA ###################################
res.condes<-condes(df,3)

res.condes$quanti  # Global association to numeric variables
res.condes$quali # Global association to factors
res.condes$category  # Partial association to significative levels in factors



 

library(FactoMineR)
summary(df$y.bin)
# The "variable to describe cannot have NA ###################################
res.catdes<-catdes(df,13)

res.catdes$quanti.var  # Global association to numeric variables
res.catdes$quanti # Partial association of numeric variables to levels of outcome factor
res.catdes$test.chi2 # Global association to factors
res.catdes$category  # Partial association to significative levels in factors



```



### Definition of binary outcome: Audi

Create binary target, define lists of numeric and qualitative variables and save your raw base database

```{r}

# Binary Target: Audi?

df$Audi<-ifelse(df$manufacturer == "Audi",1,0)##1 es compleix
df$Audi<-factor(df$Audi,labels=paste("Audi",c("No","Yes")))#treure Audi verd
summary(df$Audi)
# Pie
piepercent<-round(100*(table(df$Audi)/nrow(df)),dig=2); piepercent
pie(table(df$Audi),col=heat.colors(2),labels=paste(piepercent,"%"))
legend("topright", levels(df$Audi), cex = 0.8, fill = heat.colors(2))
# Bar Chart
barplot(table(df$Audi),main="Barplot Binary Outcome - Factor",col=c("red","green"))

```












