---
title: "Load Data and Set Sample"
author: "Josep Franquet Fabregas"
date: \today
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    toc: no
    toc_depth: '4'
  word_document:
    toc: no
    toc_depth: '4'
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
subtitle: 'Laboratori 1 - Data Preparation'
classoption: a4paper
editor_options: 
  chunk_output_type: inline
---

# USED CAR PRICES CASE STUDY -- DELIVERABLE 1

## Data Description: 100,000 UK Used Car Data set

This data dictionary describes data (<https://www.kaggle.com/adityadesai13/used-car-dataset-ford-and-mercedes>) - A sample of 5000 trips has been randomly selected from Mercedes, BMW, Volkwagen and Audi manufacturers. So, firstly you have to combine used car from the 4 manufacturers into 1 dataframe.

The cars with engine size 0 are in fact electric cars, nevertheless Mercedes C class, and other given cars are not electric cars,so data imputation is requered.

-   manufacturer Factor: Audi, BMW, Mercedes or Volkswagen
-   model Car model
-   year registration year
-   price price in £
-   transmission type of gearbox
-   mileage distance used
-   fuelType engine fuel
-   tax road tax
-   mpg Consumption in miles per gallon\
-   engineSize size in litres

## Load Required Packages: to be increased over the course

```{r}
# Load Required Packages: to be increased over the course
options(contrasts=c("contr.treatment","contr.treatment"))

requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","dplyr","ggmap","ggthemes","knitr")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(requiredPackages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
#verify they are loaded
search()

```

## Load our sample

```{r}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())

setwd("C:/Users/Miquel/Documents/GitHub/ADEI")
filepath<-"C:/Users/Miquel/Documents/GitHub/ADEI/"

load(paste0(filepath,"MyOldCars-Raw.RData"))
```

## Some useful functions

```{r, echo=FALSE}

# Mout <- which((df$tax < var_out$mouti)|(df$tax > var_out$mouts))

# Some useful functions
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

```

## Validation of the Data Set: description of the process

### Univariate Descriptive Analysis 

#### Model variable
```{r}

df$model<-factor(paste0(df$manufacturer,"-",df$model))
df$model<-factor(df$model,labels=paste0("f.Model",levels(df$model)))

summary(df$model)
barplot(summary(df$model), main = "Model Barplot", col = "blue")

```

#### Year variable
```{r}

```

#### Transmission variable
```{r}

```

#### fuelType variable
```{r}

```

#### engineSize variable
```{r}

```

#### manufacturer variable
```{r}

```

#### price variable
```{r}
df$f.price<-factor(cut(df$price/1000,breaks=c(quantile(df$price,seq(0,1,0.25),na.rm=TRUE)/1000),include.lowest = T ))
levels(df$f.price)<-paste("f.price-",levels(df$f.price))
summary(df$price)
barplot(summary(df$price), main = "Price Barplot", col = "blue")


```

#### mileage variable
```{r}

```

### Data Quality Report
#### engineSize
```{r}
sel<-which(df$engineSize==0 & df$fuelType!="f.Fuel-Electric")
ierrs[sel]<-ierrs[sel]+1
jerrs[9] <- length(sel)


```












